Fixing Vulnerabilities Potentially Hinders Maintainability

### 2020-08-21 paper/emse20.tex
update

>>>>

\section{Plan A}

Please, pay more attention to the cases with !.

\begin{itemize}
	\item How this is a problem or has an impact to the software project? 
  What is the message to developers? (clarify)
	\item Implications: Discuss that depending on the project, maybe security 
  is a first class citizen and should be fixed above all. Don't 
  just state that patch risk assessment should take maintainability into account. 
  Saying that security should be a priority requires maybe a user study
  as evidence (check if the literature available is not enough).
  \item Review the secure programming languages implication. Check 
  if the idea of fully secure exists. The idea should be more secure, not
  fully secure.
  \item [!] \textit{I am not sure how the computer science curricula got mentioned 
    from a motivation. I don't think vulnerabilities are been introduced, its 
    more that they get exposed. In this case, the implication does not make 
    much sense from a practical point of view.} -> Why not? Shouldn't 
    maintainable security be introduced in the computer science curricula since
    we have evidence that developers might not following good practices while
    patching? If they learn it in college and are aware of it in their jobs
    won't they be better prepared?
  \item [!] \textit{It did take me several readings to understand the paper, especially the approach} - 
  what to do with this @rui?
  \item [!] \textit{Fig 4, although it looks impressive, I am not sure how to read the figure} - what 
  should we do @rui?
  \item [!] \textit{Fig 5, 6 and 7 are not the best ways to portray your findings. For instance, I am 
  still not sure how the positive, none and negative metrics are calculated and normalized} - what
  is the best way then? 
  \item \textit{Authors carefully highlight the key messages of each research question and answer
  all accordingly} - Improve this.
  \item [!] \textit{wrong assumption: patching commit should also be responsible to preserve the 
  maintainability.... The patching commit is to fix the vulnerability and to put the system is 
  an error-free, executable state so that the fix can be tested} -> I agree with this for vulnerabilities
  that are the result of errors. For example, a bad expection handling, a memory leak that led to 
  a denial-of-service attack, NPE, etc. However, there are many other vulnerabilities whose
  exploits do not have interference on systems execution. So, what to do here? Should we divide
  the study by vulnerabilities and their implications on systems? Should we divide it by severity levels?
  \item \textit{authors write "...important to write maintainable code to support the production
  of more secure software...". Relationship is not clear nor is it backed by some evidence} -> 
  Improve this.
  \item Drop the word refactoring. Security patches are not refactorings. Refactorings is perfective
  or preventive maintenance, whereas patching a vulnerability is corrective maintenance.
  \item Discuss better the fact of not using other important aspects for the baseline such as the
  size of the changed method and complexity by metrics. Soften the part where it says we should
  have gone beyond size.
  \item \textit{ Figure 4 does not indicate that or support the statement that “patching 
  vulnerabilities has a very significant number of negative cases for all guidelines”. Besides a 
  corresponding statistical test that supports this statement, even the swarm and box plotz itself 
  clearly tell a different story. So, authors may rethink this statement or better explain how they 
  come to this conclusion.} -> We are overselling here. Improve it.
  \item \textit{Authors state that they assume for the Pontas dataset that the last commit related 
  to patching being the actual patch while the parent of the first patch-related commit is assumed 
  to be the vulnerable version. However, nothing is said whether this assumption holds all the time. 
  So, how did authors validate their assumption? This is missing, but should be added.} - Improve this.
  \item \textit{In Section 3-C, for computing and comparing the maintainability, only security commits  
  (and parent commit) are mentioned. Did you apply the same method also the the regular changes that 
  you use to compare with security commits? Please make this clear, currently this is ignored in the 
  particular  (sub) section.} - Improve this.
  \item \textit{For clarification, the authors explained that they performed a manual investigation 
  for RQ3. It would be beneficial for more details.} -> Clarify this.
  \item \textit{In terms of presentation, the results were a bit difficult to follow at times as they 
  did not contain contextualization. For example, “negative cases” can seem a bit ambiguous when 
  referencing the relationship between the impact with guidelines.} -> Improve this.

\end{itemize}

\section{Plan B}

Baseado nos reviewers, eu acho que para seguir as orientaçoes deles de refletir e fortalecer 
a nova assumption e as implicaçoes, teriamos de acrescentar ao Plano A, os proximos pontos
antes de submeter à ESME.

\begin{itemize}
\item \textbf{Step 1:} Pegar nos casos em que o impacto na maintainability foi negativa e ver se foram modificados 
depois e se sim, se houve melhoria da maintainability. -> isto para então resolver a questao da 
assumption errada.
\item \textbf{Step 2:} Se não existir nenhuma alteração, entao talvez entrar em contacto com os developers
e perguntar se quando fizeram o patch tiveram em consideração maintainability do projecto. E se não porquê.
\item Problemas de baselines e testes acho que nao vale a pena pegar porque vai depender
do reviewer (eu acho).
\end{itemize}

Qual é o plano que acha melhor? Como prosseguir?

\section{Later}

\begin{itemize}
	\item \textbf{Assumptions:}\textit{authors measure the maintainability directly after the patch. inestad it should look 
  for multiple follow up commits and how maintainability evolves there, as it is likely that certain
  issues are addresses AFTER the patching commit} -> I think we should see if later commits touch on
  the patch and measure their maintainability again. Was a less maintainable fix pushed first and then
  refactored, or does the code stay less maintainable?
  \item \textbf{To back up implications:} \textit{In terms of the push for changes to curricula, I would agree that it is important 
  to always stress writing high quality code. However, the support by the study is not strong. Given a 
  majority of patches did not decrease maintainability, it shows that there is a reasonable focus on maintainability. 
  Similarly, it is important to understand the circumstances when it decreased. As previously mentioned, organizations 
  like would prioritize getting the system patched faster (especially in the case of a severe vulnerability) over 
  its maintainability.} -> Work with practioners to understand if they were taking into consideration maintainability?
  \item \textbf{Baseline:}\textit{Authors just search for commits of a similar size. There are other important aspects
  that are entirely neglected such as size of the changed method or how complex it it (by means of 
  metrics)} -> Improve baseline.
  \item \textbf{Baseline:}\textit{Similarly, the comparison to “regular” commits really compares security patches to 
  any other type of change. It would have been more meaningful to compared them to specific types of 
  commits, like bug fixes (e.g., are security fixes less maintainable than bug fixes?).}
  \item \textbf{Tests:}\textit{or the chosen statistical test, the hypothesis may be not well chosen as it just checks for 
  the same distribution of pairs of versions. However, this means if a similar number of pairs have 
  a negative and a positive impact, the hypothesis holds however, if this is not the case, the statistical 
  test does not state in which direction the hypothesis is rejected, that is, what eventually is the 
  alternative hypothesis. But this is necessary to get insights whether maintainability is hampered or not. 
  Authors may rethink the choice of the hypothesis or even the test itself.} -> What about this?
  \item \textbf{Thesis idea (PL):} It is important 
  to understand the correlation between the severity of the vulnerability and the programming language as well 
  as the maintainability impact. It is not clear how software written a more security-focused language would 
  inherently be more maintainable (it assumes the flaw is the language and not the person).  
\end{itemize}
